---
output: 
  html_document:
      css: styles.css
runtime: shiny
---
  
# OC MS4 monitoring program, dry weather monitoring {.tabset}
  
```{r setup, message = F, warning = F, results = 'hide', echo = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, echo = F, fig.path = 'figs/', dev.args = list(family = 'serif'))

library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(lubridate)
library(leafsync)
library(viridisLite)
library(lubridate)
library(gridExtra)
library(stargazer)
library(EnvStats)
library(shiny)
library(kableExtra)
library(mgcv)
library(metR)
library(shinyWidgets)
library(plotly)
library(english)
library(vegan)
library(ggord)
library(patchwork)

mptyps <- c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "Esri.WorldImagery", "OpenTopoMap")

mapviewOptions(leafletHeight = 300)

prj <- 4326 # wgs84

source('R/funcs.R')

##
# ggplot themes

thm1 <- theme_bw(base_size = 12) + 
  theme(
    strip.background = element_blank(), 
    strip.placement = 'outside', 
    axis.title.x = element_blank(), 
    legend.title = element_blank(), 
    legend.position = 'bottom', 
    panel.grid = element_blank()
  )

thm2 <- theme_bw(base_size = 16) + 
  theme(
    strip.background = element_blank(), 
    strip.placement = 'outside', 
    legend.position = 'bottom', 
    panel.grid = element_blank()
  )

pbase <- theme_bw(base_family = 'serif') +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12), 
    axis.text.y = element_text(size = 12),
    legend.position = 'top',
    legend.direction = 'horizontal',
    # plot.margin = unit(c(4,4,0,0), "lines"),
    strip.background = element_blank(), 
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
    panel.background = element_rect(fill = 'black')
  ) 

data(dwdat)
data(thrsdat)
data(tmdldat)

# xy locs
locs <- dwdat %>%
  select(StationCode, Longitude, Latitude) %>%
  unique

# date ranges
dts <- dwdat %>% 
  pull(Date) %>% 
  range

# obs ranges per station
obsrng <- dwdat %>% 
  group_by(StationCode, Parameter) %>% 
  summarise(n = n()) %>% 
  pull(n) %>% 
  range

# obs ranges per station, averages across parameters
obsave <- dwdat %>% 
  group_by(StationCode, Parameter) %>% 
  summarise(n = n()) %>% 
  group_by(StationCode) %>% 
  summarise(n = mean(n, na.rm = T)) %>% 
  mutate(n = round(n, 1))
obsaverng <- obsave %>% 
  pull(n) %>% 
  range

# color palette for hotspot exceedances, as prop
hotcol <- colorNumeric(
  palette = rev(RColorBrewer::brewer.pal(11, 'RdYlBu')),
  na.color = 'yellow',
    domain = c(0, 100)
  )

# nutrients, top parameters (top ten plus those with thresholds)
nutrs <- c('Ammonia', 'Nitrate, Nitrite', 'Total Kjeldahl Nitrogen', 'Total Nitrogen', 'Orthophosphate', 'Total Phosphorus')
tops <- table(dwdat$Parameter) %>% sort %>% rev %>% .[1:10] %>% names %>% sort
tops <- thrsdat %>% 
  pull(Parameter) %>% 
  c(., tops) %>% 
  unique
tops <- tops[!tops %in% nutrs]

# lookup list for parameters with thresholds and TMDL waterbodies
tmdl <- list(
  EC = 'EC', 
  Metals = c('Ag', 'As', 'Cd', 'Cr', 'Cu', 'Hg', 'Ni', 'Pb', 'Se', 'Zn'), 
  Nutrients = c('Ammonia', 'Nitrate, Nitrite'), 
  Pathogens = c('ENT', 'FC', 'TC'),
  Pesticides = c('Chlorpyrifos', 'Diazinon', 'Dimethoate', 'Fensulfothion', 'Malathion'), 
  TSS = 'TSS'
  ) %>% 
  enframe('grp', 'Parameter') %>% 
  unnest(Parameter)
```

```{r reactives}
# tabular inventory by station, parameter
statab <- reactive({
  
  # input
  wshsel2 <- input$wshsel2
  varsel2 <- input$varsel2
  tabsel <- input$tabsel
  
  dat <- dwdat %>% 
    filter(Watershed %in% wshsel2) %>% 
    filter(Parameter %in% varsel2)
  
  if(tabsel == 'by year'){
    
    sums <- dat %>% 
      mutate(Year = as.character(year(Date))) %>% 
      group_by(Year, StationCode) %>% 
      summarise(n = n()) %>% 
      ungroup() %>%  
      bind_rows(group_by(., Year) %>%
                  summarise(n = sum(n)) %>%
                  mutate(StationCode = 'Total')) %>%
      bind_rows(group_by(., StationCode) %>%
                  summarise(n=sum(n)) %>%
                  mutate(Year = 'Total')) %>%
      spread(Year, n, fill = 0)

  }

  if(tabsel == 'by month'){
    
    sums <- dat %>% 
      mutate(Month = month(Date, label = TRUE)) %>% 
      group_by(Month, StationCode) %>% 
      summarise(n = n()) %>% 
      ungroup() %>% 
      bind_rows(group_by(., Month) %>%
                  summarise(n = sum(n)) %>%
                  mutate(StationCode = 'Total')) %>%
      bind_rows(group_by(., StationCode) %>%
                  summarise(n=sum(n)) %>%
                  mutate(Month = 'Total')) %>%
      mutate(Month = factor(Month, levels = c("May", "Jun", "Jul", "Aug", "Sep", "Total"))) %>%
      spread(Month,n,fill=0)
    
  }
  
  alltot <- sums %>% 
    filter(StationCode != 'Total')
  coltot <- sums %>% 
    filter(StationCode == 'Total')

  totab <- bind_rows(alltot, coltot)

  tab <- HTML(knitr::kable(totab, format = 'html', caption = paste(varsel2, tabsel)) %>% 
       kable_styling(full_width = T, font_size = 14))
    
  
  return(tab)
  
})

# plot inventory by station, parameter
staplo <- reactive({
  
  # input
  wshsel2 <- input$wshsel2
  varsel2 <- input$varsel2
  tabsel <- input$tabsel
  
  dat <- dwdat %>% 
    filter(Watershed %in% wshsel2)

  if(tabsel == 'by year'){
    
    colrng = c(0, 7)
    
    sums <- dat %>% 
      mutate(Year = as.character(year(Date))) %>% 
        group_by(Year, StationCode, Parameter) %>% 
        summarise(n = n()) %>% 
        ungroup() %>% 
        mutate(Year = factor(Year, levels = c(2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019))) %>% 
        complete(Year, StationCode, Parameter, fill = list(n = 0)) %>% 
        filter(Parameter %in% varsel2) %>% 
        rename(x = Year)

  }

  if(tabsel == 'by month'){
    
    colrng <- c(0, 12)
    
    sums <- dat %>% 
      mutate(Month = month(Date, label = TRUE)) %>% 
      group_by(Month, StationCode, Parameter) %>% 
      summarise(n = n()) %>% 
      ungroup() %>% 
      mutate(Month = factor(Month, levels = c("May", "Jun", "Jul", "Aug", "Sep"))) %>% 
      tidyr::complete(Month, StationCode, Parameter, fill = list(n = 0)) %>% 
      filter(Parameter %in% varsel2) %>% 
      rename(x = Month)
    
  }
  
  toplo <- sums %>% 
    mutate(StationCode = factor(StationCode, levels = rev(unique(StationCode))))
  txtplo <- toplo %>% 
    filter(n > 0)

  p <- ggplot(toplo) +
    geom_tile(aes(x = x, y = StationCode, fill = n), colour = 'black') +
    geom_text(data = txtplo, aes(x = x, y = StationCode, label = n)) + 
    scale_x_discrete('', expand = c(0, 0)) +
    scale_y_discrete('', expand = c(0, 0)) +
    scale_fill_gradientn('Number of samples', colors = RColorBrewer::brewer.pal(9, 'BuGn'), limits = colrng) +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10, label.theme = element_text(size = 11, angle = 0))) +
    pbase
  
  return(p)
  
})

# get threshold variable
varthr <- reactive({
  
  # input
  varsel3 <- input$varsel3
  
  out <- thrsdat %>% 
    filter(Parameter %in% varsel3) %>% 
    pull(Threshold)
  
  return(out)
  
})

# hotspot input ranges
hotrng <- reactive({
  
  # inputs
  varsel3 <- input$varsel3
  varthr <- varthr()
  
  # values for slider
  qnts <- c(0, 0.5, 1)
  out <- dwdat %>% 
    filter(Parameter %in% varsel3) %>%
    pull(Result) %>% 
    quantile(probs = qnts, na.rm = T) %>%
    round(2)

  # make the default selection the threshold if applicable
  if(length(varthr) != 0)
    out[2] <- varthr
    
  return(out)
  
})

# hotspot data
hotdat <- reactive({
  
  # inputs
  cntsel <- input$cntsel
  varsel3 <- input$varsel3
  dtssel <- input$dtssel
  thrsel2 <- input$thrsel2
  tmdsel <- as.logical(input$tmdsel)
  locsel <- input$locsel
  req(thrsel2)

  # filter by tmdl waterbodies
  if(tmdsel & length(locsel) > 0){
    
    grp <- tmdl %>% 
      filter(Parameter %in% varsel3) %>% 
      pull(grp) 

    stas <- tmdldat %>% 
      filter(Parameter %in% !!grp) %>% 
      filter(Receiving %in% locsel) %>% 
      select(StationCode, Receiving)

    out <- dwdat %>% 
      inner_join(stas, by = 'StationCode') %>% 
      select(-Watershed) %>% 
      rename(Watershed = Receiving)
  
  }
  
  if(!tmdsel)
    out <- dwdat 
  
  req(exists('out'))
  
  # summary for all dates, stations by selected constituent
  out <- out %>% 
    filter(Parameter %in% varsel3) %>%
    group_by(Watershed, StationCode) %>% 
    mutate(n = n()) %>% 
    filter(Date >= dtssel[1] & Date <= dtssel[2]) %>%
    mutate(
      abv = case_when(
        Result > thrsel2 ~ 1, 
        T ~ 0
      )
    ) %>% 
    summarise(
      exceeds = sum(abv),
      n = unique(n)
      ) %>% 
    ungroup() %>% 
    mutate(
      exceeds = round(100 * exceeds / n, 0), 
      cols = hotcol(exceeds), 
      cexs = scales::rescale(exceeds, to = c(4, 17), from = c(0, 100))
    ) %>% 
    select(Watershed, StationCode, exceeds, n, cols, cexs) %>% 
    filter(n >= cntsel[1] & n <= cntsel[2]) %>% 
    unique
  
  return(out)
  
})

# map of exceedances
hotmap <- reactive({
  
  # input
  hotdat <- hotdat()
  
  req(nrow(hotdat) > 0)
  
  tomap <- hotdat %>%
    left_join(locs, by = 'StationCode') %>%
    st_as_sf(coords = c("Longitude", "Latitude"), crs = prj)

  # hover pt labels
  labs <- paste(tomap$StationCode, ': ', tomap$exceeds, ' % exceeding, ', tomap$n, ' total obs.')
  leglab <- "% exceeding"
  
  mapviewOptions(leafletHeight = 400)

  m <- mapview(tomap, cex = tomap$cexs, label = labs, legend = F, layer.name = F, col.regions = tomap$cols, homebutton = F, map.types = mptyps)
  
  # add legend
  out <- m@map %>% 
    addLegend("bottomright", pal = hotcol, title = leglab, opacity = 1, values = tomap$exceeds) 
  
  return(out)
  
})

# plot of hotspots by watershed
hotplo <- reactive({
  
  # input
  hotdat <- hotdat()
  dtssel <- input$dtssel
  
  req(nrow(hotdat) > 0)
  
  toplo <- hotdat %>% 
    arrange(Watershed, exceeds) %>% 
    mutate(StationCode = factor(StationCode, levels = .$StationCode))

  ylab <- paste0("Number of exceedances above the median\nfrom ", as.character(dtssel[1]), ' to ', as.character(dtssel[2]))
  
  p <- ggplot(toplo, aes(x = StationCode, y = exceeds)) + 
    geom_bar(stat = 'identity', fill = toplo$cols, color = 'grey') + 
    facet_grid(~ Watershed, scales = 'free_x', space = 'free_x') + 
    thm1 + 
    theme(
      axis.text.x = element_text(size = 7, angle = 45, hjust = 1)
    ) + 
    scale_y_continuous(limits = c(0, 100)) + 
    labs(
      y = ylab
    )

  return(p)
  
})

# quantiles based on thrsel3
hotrng2 <- reactive({
  
  # input
  thrsel3 <- input$thrsel3

  out <- dwdat %>% 
    filter(Parameter %in% c(tops, nutrs)) %>%
    group_by(Parameter) %>% 
    summarise(
      qnt = quantile(Result, probs = as.numeric(thrsel3), na.rm= T)
    ) %>% 
    left_join(thrsdat, by = 'Parameter') %>% 
    mutate(
      qnt = case_when(
        is.na(Threshold) ~ qnt, 
        T ~ Threshold
      )
    )
  
  return(out)
  
})

# hotspot data
hotdat2 <- reactive({
  
  # inputs
  cntsel2 <- input$cntsel2
  dtssel2 <- input$dtssel2
  hotrng2 <- hotrng2()
  
  req(hotrng2)
  
  # summary for all dates, stations by selected constituent
  out <- dwdat %>% 
    filter(Parameter %in% c(tops, nutrs)) %>%
    left_join(hotrng2, by = 'Parameter') %>% 
    left_join(obsave, by = 'StationCode') %>% 
    filter(n >= cntsel2[1] & n <= cntsel2[2]) %>% 
    filter(Date >= dtssel2[1] & Date <= dtssel2[2]) %>%
    mutate(
      abv = case_when(
        Result > qnt ~ 1, 
        T ~ 0
      )
    ) %>% 
    group_by(StationCode, Watershed, Parameter, n) %>% 
    summarise(
      exceeds = sum(abv, na.rm = T)
      ) %>% 
    group_by(StationCode, Watershed, n) %>% 
    summarise(
      exceeds = mean(exceeds, na.rm = T)
    ) %>% 
    ungroup() %>% 
    mutate(
      exceeds = round(100 * exceeds / n, 0), 
      cols = hotcol(exceeds), 
      cexs = scales::rescale(exceeds, to = c(4, 17), from = c(0, 100))
    ) %>% 
    select(Watershed, StationCode, exceeds, n, cols, cexs) %>% 
    unique
  
  return(out)
  
})

# map of exceedances
hotmap2 <- reactive({
  
  # input
  hotdat2 <- hotdat2()
  
  tomap <- hotdat2 %>%
    left_join(locs, by = 'StationCode') %>%
    st_as_sf(coords = c("Longitude", "Latitude"), crs = prj)

  # hover pt labels
  labs <- paste(tomap$StationCode, ': ', tomap$exceeds, ' avg % exceedences, ', tomap$n, ' total avg obs.')
  leglab <- "% exceedances"
  
  mapviewOptions(leafletHeight = 400)

  m <- mapview(tomap, zcol = 'n', cex = tomap$cexs, label = labs, legend = F, layer.name = F, col.regions = tomap$cols, homebutton = F, map.types = mptyps)
  
  # add legend
  out <- m@map %>% 
    addLegend("bottomright", pal = hotcol, title = leglab, opacity = 1, values = tomap$exceeds) 
  
  return(out)
  
})

# plot of hotspots by watershed
hotplo2 <- reactive({
  
  # input
  hotdat2 <- hotdat2()
  dtssel <- input$dtssel
  
  toplo <- hotdat2 %>% 
    arrange(Watershed, exceeds) %>% 
    mutate(StationCode = factor(StationCode, levels = .$StationCode))

  ylab <- paste0("Number of exceedances above the median\nfrom ", as.character(dtssel[1]), ' to ', as.character(dtssel[2]))
  
  p <- ggplot(toplo, aes(x = StationCode, y = exceeds)) + 
    geom_bar(stat = 'identity', fill = toplo$cols, color = 'grey') + 
    facet_grid(~ Watershed, scales = 'free_x', space = 'free_x') + 
    thm1 + 
    theme(
      axis.text.x = element_text(size = 7, angle = 45, hjust = 1)
    ) + 
    scale_y_continuous(limits = c(0, 100)) + 
    labs(
      y = ylab
    )

  return(p)
  
})

# get threshold variable for watershed plots
varthr2 <- reactive({
  
  # input
  varsel <- input$varsel
  
  out <- thrsdat %>% 
    filter(Parameter %in% varsel) %>% 
    pull(Threshold)
  
  return(out)
  
})

# hotspot input ranges for watershed analysis
hotrng3 <- reactive({
  
  # inputs
  varsel <- input$varsel
  varthr2 <- varthr2()
  
  # values for slider
  qnts <- c(0, 0.5, 1)
  out <- dwdat %>% 
    filter(Parameter %in% varsel) %>%
    pull(Result) %>% 
    quantile(probs = qnts, na.rm = T) %>%
    round(2)

  # make the default selection the threshold if applicable
  if(length(varthr2) != 0)
    out[2] <- varthr2
    
  return(out)
  
})

# watershed and parameter data for observed plots
wshpardat <- reactive({

  # inputs
  locsel2 <- input$locsel2
  varsel <- input$varsel
  tmdsel2 <- as.logical(input$tmdsel2)
  
  req(locsel2)
  
  if(tmdsel2){
    
    grp <- tmdl %>% 
      filter(Parameter %in% varsel) %>% 
      pull(grp)
    stas <- tmdldat %>% 
      filter(Parameter %in% !!grp) %>% 
      filter(Receiving %in% locsel2) %>% 
      pull(StationCode)
    
    out <- dwdat %>% 
      filter(StationCode %in% !!stas)

  }
  
  if(!tmdsel2)
    out <- dwdat %>%
      filter(Watershed %in% locsel2)
  
  req(exists('out'))
  
  out <- out %>% 
    mutate(
      Year = year(Date),
      Month = month(Date, label = T)
    ) %>%
    filter(Parameter %in% varsel)

  return(out)

})

# further subset of wshpardat by stations
stawshpardat <- reactive({
  
  # inputs
  wshpardat <- wshpardat()
  stasel <- input$stasel

  out <- wshpardat %>% 
    filter(StationCode %in% stasel)
  
  return(out)
  
})

# stawshspardat with threshold
thrstawshpardat <- reactive({
  
  req(nrow(stawshpardat()) > 0)
  req(!is.null(input$thrsel))
  
  # inputs
  stawshpardat <- stawshpardat()
  thrsel <- as.numeric(input$thrsel)

  out <- stawshpardat %>% 
    mutate(
      abvthr = case_when(
        Result >= thrsel ~ 'Above threshold', 
        Result < thrsel ~ 'Below threshold'
      ), 
      abvthr = factor(abvthr, levels = c('Above threshold', 'Below threshold'))
    )
  
  return(out)
  
})

# text summary of thresholds analyses
thrtx <- reactive({
  
  req(nrow(thrstawshpardat()) > 0)
  
  # input
  thrstawshpardat <- thrstawshpardat()
  thrsel <- as.numeric(input$thrsel)
  
  # n stations exceeding
  nsta <- thrstawshpardat %>% 
    group_by(StationCode) %>% 
    summarise(abv = any(abvthr == 'Above threshold')) %>% 
    pull(abv) %>% 
    sum
  
  # total stations
  totsta <- thrstawshpardat %>% 
    pull(StationCode) %>% 
    unique %>% 
    length
  
  # percent stations exceeding
  persta <- round(100 * nsta / totsta, 1)
  
  # nobs exceeding
  nobs<- thrstawshpardat %>% 
    mutate(abv = abvthr == 'Above threshold') %>% 
    pull(abv) %>% 
    sum
  
  # total obs
  totobs <- nrow(thrstawshpardat)
  
  # percent observations exceeding
  perobs <- round(100 * nobs / totobs, 1)
  
  out <- paste0('For the threshold ', thrsel, ': ', english(nsta), ' of ', english(totsta), ' stations (', persta, '%) exceeding, ', nobs, ' observations of ', totobs, ' exceeding (', perobs, '%).')
  
  return(out)
  
})

# first observed plot
obsp1 <- reactive({

  req(nrow(thrstawshpardat()) > 0)
  
  # inputs
  thrstawshpardat <- thrstawshpardat()
  wshsel <- input$wshsel
  logsel <- input$logsel
  thrsel <- as.numeric(input$thrsel)

  toplo <- thrstawshpardat
  
  # labels
  ylb <- paste0('Concentration (', unique(toplo$Units), ')')
  ttl <- unique(toplo$Parameter)

  p <- ggplot(toplo, aes(x = Date, y = Result, group = StationCode)) +
    geom_hline(yintercept = thrsel, linetype = 'dotted', colour = 'tomato1') +
    geom_line() +
    geom_point(aes(colour = abvthr), alpha = 0.6) +
    scale_colour_manual(values = c('tomato1', 'black'), drop = F) +
    thm1 +
    labs(
      subtitle = wshsel,
      title = ttl,
      y = ylb
    )
  
  if(logsel)
    p <- p + 
      scale_y_log10()

  p <- ggplotly(p)
  
  return(p)

})

# second observed plot
obsp2 <- reactive({

  req(nrow(thrstawshpardat()) > 0)
  
  # inputs
  thrstawshpardat <- thrstawshpardat()
  wshsel <- input$wshsel
  logsel <- input$logsel
  thrsel <- as.numeric(input$thrsel)

  toplo <- thrstawshpardat
  
  # labels
  ttl <- paste(unique(toplo$Parameter), 'by year')
  ylb <- paste0('Concentration (', unique(toplo$Units), ')')

  p <- ggplot(toplo, aes(x = factor(Year), y = Result, group = Year, label = StationCode)) +
    geom_hline(yintercept = thrsel, linetype = 'dotted', colour = 'tomato1') +
    geom_boxplot(outlier.shape = NA) +
    geom_point(aes(colour = abvthr), position = position_jitter(width = 0.2), alpha = 0.6) +
    scale_colour_manual(values = c('tomato1', 'black'), drop = F) +
    thm1 +
    labs(
      subtitle = wshsel,
      title = ttl,
      y = ylb
    )
  
  if(logsel)
    p <- p + 
      scale_y_log10()

  p <- ggplotly(p)
  
  # remove outliers 
  p$x$data[[2]]$marker = list(opacity = 0) 
  
  return(p)

})

# third observed plot
obsp3 <- reactive({

  req(nrow(thrstawshpardat()) > 0)
  
  # inputs
  thrstawshpardat <- thrstawshpardat()
  wshsel <- input$wshsel
  logsel <- input$logsel
  thrsel <- as.numeric(input$thrsel)
  
  toplo <- thrstawshpardat
  
  # labels
  ttl <- paste(unique(toplo$Parameter), 'by month')
  ylb <- paste0('Concentration (', unique(toplo$Units), ')')
  
  p <- ggplot(toplo, aes(x = Month, y = Result, group = Month, label = StationCode)) +
    geom_hline(yintercept = thrsel, linetype = 'dotted', colour = 'tomato1') +
    geom_boxplot(outlier.shape = NA) +
    geom_point(aes(colour = abvthr), position = position_jitter(width = 0.2), alpha = 0.6) +
    scale_colour_manual(values = c('tomato1', 'black'), drop = F) +
    thm1 +
    labs(
      subtitle = wshsel,
      title = ttl,
      y = ylb
    )

  if(logsel)
    p <- p + 
      scale_y_log10()
  
  p <- ggplotly(p)
  
  # remove outliers 
  p$x$data[[2]]$marker = list(opacity = 0) 

  return(p)

})
```

```{r downloads}
output$downloadData <- downloadHandler(
  filename = function() {"dry_weather.csv"},
  content = function(file) {
    write.csv(dwdat, file, row.names = FALSE)
  }
)
```

This website presents an evaluation of the Orange County dry weather (ID/IC) dataset. The content is separated into three main tabs.  

* __Inventory__: Map-based and tabular summaries of monitoring effort and basic characteristics of the data
* __Hotspots__: Assessment of hotspot sites based on threshold exceedances over time
* __Analyses by waterbody__: A simple analysis of threshold exceedances for select sites shown for complete time series

Each main tab includes sub-tabs or drop-down menus for selecting and viewing different content.  Selected constituents are also available for each analysis.  Because the constituent list was large for some stations, a pre-selected list was created that included the top ten measured constituents for the entire data set and those those with defined thresholds.   Nutrients were also added to the list of constituents available for analysis.  The following shows which constituents were included in the website. 

* __Top ten, including those with thresholds__: `r paste(tops, collapse = ', ')`
* __Nutrients__: `r paste(nutrs, collapse = ', ')`

## Inventory {.tabset .tabset-pills}

```{r}
downloadBttn('downloadData', 'Download data', style = 'simple', block = F, color = 'success', size = 'sm')
```
<br>

### Maps

These three maps show the stations for the dry weather monitoring program.  Each map shows different information, the first showing the watershed for each station, the second showing the number of parameters collected at each station (for the entire period of record), and the third showing the number of years for available data.  Each map is linked so that the mouse pointer is synced between the three.

```{r}
tomap <- dwdat %>% 
  select(StationCode, Watershed, Longitude, Latitude) %>% 
  unique %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = prj)

m1a <- mapview(tomap, zcol = 'Watershed', layer.name = 'Watershed location', homebutton = F, map.types = mptyps)

tomap <- dwdat %>% 
  select(StationCode, Parameter, Longitude, Latitude) %>% 
  unique %>% 
  group_by(StationCode, Longitude, Latitude) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = prj)

m2a <- mapview(tomap, zcol = 'n', layer.name = 'Number of parameters', homebutton = F, col.regions = magma, map.types = mptyps)

tomap <- dwdat %>% 
  select(StationCode, yr = Date, Longitude, Latitude) %>% 
  mutate(yr = year(yr)) %>% 
  unique %>% 
  group_by(StationCode, Longitude, Latitude) %>% 
  summarise(n = n()) %>% 
  ungroup %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = prj)

m3a <- mapview(tomap, zcol = 'n', layer.name = 'Number of years', homebutton = F, col.regions = magma, map.types = mptyps)

leafsync::sync(m1a, m2a, m3a, ncol = 1) 
```

### Tables, by watershed

These tables summarize the number of observations for each constituent by each major watershed.  The first three tables show the summaries for the top parameters and the bottom three show the summaries for nutrients.  Tables are also shown for total count, number of observations within each year, and number of observations within each month for the top parameters and nutrients. 

```{r toptab}
sums <- dwdat %>% 
  filter(Parameter %in% tops) %>% 
  group_by(Parameter, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Parameter) %>%
              summarise(n=sum(n)) %>%
              mutate(Watershed='Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Parameter='Total')) %>%
  spread(Parameter,n,fill=0) %>% 
  select_at(c('Watershed', tops, 'Total'))

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Top parameters") %>% 
       kable_styling(full_width = T, font_size = 14))
```

```{r toptabyr}
sums <- dwdat %>% 
  mutate(Year = as.character(year(Date))) %>% 
  filter(Parameter %in% tops) %>% 
  group_by(Year, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Year) %>%
              summarise(n = sum(n)) %>%
              mutate(Watershed = 'Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Year = 'Total')) %>%
  spread(Year,n,fill=0)

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Top parameters by year") %>% 
       kable_styling(full_width = T, font_size = 14))
```

```{r toptabmo}
sums <- dwdat %>% 
  mutate(Month = month(Date, label = TRUE)) %>% 
  filter(Parameter %in% tops) %>%
  group_by(Month, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Month) %>%
              summarise(n = sum(n)) %>%
              mutate(Watershed = 'Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Month = 'Total')) %>%
  mutate(Month = factor(Month, levels = c("May", "Jun", "Jul", "Aug", "Sep", "Total"))) %>%
  spread(Month,n,fill=0)

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Top parameters by month") %>% 
       kable_styling(full_width = T, font_size = 14))
```

```{r nuttab}
sums <- dwdat %>% 
  filter(Parameter %in% nutrs) %>% 
  group_by(Parameter, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Parameter) %>%
              summarise(n=sum(n)) %>%
              mutate(Watershed='Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Parameter='Total')) %>%
  spread(Parameter,n,fill=0) %>% 
  select_at(c('Watershed', nutrs, 'Total'))

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Nutrients") %>% 
       kable_styling(full_width = T, font_size = 14))
```

```{r nuttabyr}
sums <- dwdat %>% 
  mutate(Year = as.character(year(Date))) %>% 
  filter(Parameter %in% nutrs) %>% 
  group_by(Year, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Year) %>%
              summarise(n = sum(n)) %>%
              mutate(Watershed = 'Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Year = 'Total')) %>%
  spread(Year,n,fill=0)

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Nutrients by year") %>% 
       kable_styling(full_width = T, font_size = 14))
```

```{r nuttabmo}
sums <- dwdat %>% 
  mutate(Month = month(Date, label = T)) %>% 
  filter(Parameter %in% nutrs) %>% 
  group_by(Month, Watershed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  bind_rows(group_by(., Month) %>%
              summarise(n = sum(n)) %>%
              mutate(Watershed = 'Total')) %>%
  bind_rows(group_by(., Watershed) %>%
              summarise(n=sum(n)) %>%
              mutate(Month = 'Total')) %>%
  mutate(Month = factor(Month, levels = c("May", "Jun", "Jul", "Aug", "Sep", "Total"))) %>% 
  spread(Month,n,fill=0)

alltot <- sums %>% 
  filter(Watershed != 'Total')
coltot <- sums %>% 
  filter(Watershed == 'Total')

totab <- bind_rows(alltot, coltot)

HTML(knitr::kable(totab, format = 'html', caption = "Nutrients by month") %>% 
       kable_styling(full_width = T, font_size = 14))
```

### By Site

This table summarizes the number of observations for `r renderText(input$varsel2)` at stations within the `r renderText(input$wshsel2)` watershed.  The summaries can be shown by year or within months.  The plot option shows the same information as in the table, except each cell is shaded by relative number of observations.  

```{r}
column(12, 
       column(4, 
              selectInput('wshsel2', 'Select watershed:', choices = unique(dwdat$Watershed))
       ),
       column(4, 
              selectInput('varsel2', 'Select constituent:', choices = sort(c(tops, nutrs)), selected = 'ENT')
              ),
       column(4, 
              selectInput('tabsel', 'Select table:', choices = c('by year', 'by month'))
       )
)
```


#### {.tabset .tabset-pills}

##### Table 

```{r}
renderUI({statab()})
```

##### Plot 

```{r}
renderPlot({staplo()}, height = 500, width = 700)
```

## Hotspots {.tabset .tabset-pills}

This tab show hotspots for stations that have many exceedances over the period of record.  Hotspots are defined by the number of instances when the observed data were above the threshold for the constituent, divided by the total number of observations at a site.  Sites with larger points colored in red have many exceedances, as compared to sites with small points in blue that have few exceedances.  Thresholds are defined targets for each constituent and were obtained from Orange County.  Median values are applied in the absence of a threshold, which applied only to TKN, TP, and TN.

### Individual constituents

```{r}
column(12, 
       column(4, 
              selectInput('varsel3', 'Select constituent:', choices = sort(c(tops, nutrs)), selected = 'ENT')
       ),
       column(8, 
              sliderInput('dtssel', 'Select date range:', min = dts[1], max = dts[2], value = dts, timeFormat="%Y-%m", width = '600px')
       )
)
column(12,
       column(4,
              selectInput('tmdsel', 'Filter by TMDL receiving waterbodies?', choices = c(F, T))
       ),
       column(8,
              sliderInput('cntsel', 'Filter stations by total n:', min = 0, max = obsrng[2], value = c(0, obsrng[2]), width = '600px')
       )
)
column(12,
       column(4,               
              renderUI({
                
                # input
                varsel3 <- input$varsel3
                tmdsel <- as.logical(input$tmdsel)
                
                req(tmdsel)
                
                # get receiving tmdl waterbodies 
                grp <- tmdl %>% 
                  filter(Parameter %in% varsel3) %>% 
                  pull(grp)
                locs <- tmdldat %>% 
                  filter(Parameter %in% !!grp) %>% 
                  pull(Receiving) %>% 
                  unique
                
                if(length(locs) == 0)
                  out <- 'No TMDLs for selection'
                
                if(length(locs) > 0)
                  out <- pickerInput(inputId = "locsel", label = 'Select TMDL receiving waterbodies:', choices = locs,
                    options = list(`actions-box` = TRUE, size = 20), selected = locs, multiple = TRUE)      
                
                return(out)
                
              })
       ),
       column(4,
              renderUI({
                
                # input
                hotrng <- hotrng()
                
                numericInput('thrsel2', 'Input threshold:', min = hotrng[1], max = hotrng[3], value = hotrng[2])
                
              })
       ), 
       column(4, 
              renderText({
         
                # input
                hotrng <- hotrng()
                varthr <- varthr()
                
                txt <- paste0("min ", hotrng[1], ', max ', hotrng[3])
     
                if(length(varthr) == 0)
                  out <- paste0(txt, ', default selection is median (no threshold)')
                
                if(length(varthr) != 0)
                  out <- paste0(txt, ', default selection is defined threshold')
                  
                return(out)
                       
              })
       )
)
```

This map shows the percentage of observations above a threshold for `r renderText({input$varsel3})` for the period of record between `r renderText(as.character(input$dtssel[1]))` and `r renderText(as.character(input$dtssel[2]))`.  Sites are also filtered based on the number of total observations, where sites sites with as few as `r renderText(input$cntsel[1])` and as many as `r renderText(input$cntsel[2])` observations are shown below.  The threshold defining exceedances for `r renderText({input$varsel3})` is shown above.  

Sites can also be filtered based on the receiving waterbody with a TMDL.  Selecting "TRUE" from the drop-down list will create an additional list to filter the stations by the receiving waterbody. By default, all stations are shown.

```{r}
renderLeaflet(hotmap())
```

This plot has the same information as the map, but shows relative exceedances grouped by major watersheds.  Note that these watersheds may be different than the receiving TMDL waterbody selection. 

```{r}
renderPlot(hotplo(), height = 450, width = 900)
```

### Multiple constituents

```{r}
column(12, 
       column(4, NULL),
       column(8, 
              sliderInput('dtssel2', 'Select date range:', min = dts[1], max = dts[2], value = dts, timeFormat="%Y-%m", width = '600px')
       )
)
column(12,
       column(4, NULL),
       column(8,
              sliderInput('cntsel2', 'Filter stations by average n:', min = 0, max = obsaverng[2], value = c(0, obsaverng[2]), width = '600px')
       )
)
column(12,
       column(4, NULL),
       column(4,
              selectInput('thrsel3', 'Input threshold as quantile:', choices = seq(0.1, 0.9, by = 0.1), selected = 0.5)
       )
)
```

This map shows the percentage of observations above a threshold for all constituents for the period of record between `r renderText(as.character(input$dtssel2[1]))` and `r renderText(as.character(input$dtssel2[2]))`.  Sites sites with as few as `r renderText(input$cntsel2[1])` and as many as `r renderText(input$cntsel2[2])` average number of observations are shown below.  The map differs from the one in the previous sub-tab by showing the average number of exceedances across all constituents.  The exceedances are defined by quantile distribution of observations for each constituent, and in this case is set at `r renderText(round(100 * as.numeric(input$thrsel3), 0))`%.  

```{r}
renderLeaflet(hotmap2())
```

This plot has the same information as the map, but shows relative exceedances grouped by major watersheds.  Note that these watersheds may be different than the receiving TMDL waterbody selection. 

```{r}
renderPlot(hotplo2(), height = 450, width = 900)
```

## Analyses by waterbody {.tabset .tabset-pills}

This tab shows results across complete time series of exceedances relative to a threshold.  Points in red are above the threshold shown by the dashed line. The top plot shows the complete time series with lines joining observations for the same site. The middle plot shows the same values grouped by year and the bottom plot shows the same values grouped by month.  Sites can be filtered by receiving waterbodies with TMDLs or by major watershed (default). Individual sites can be also removed or added to the plots from the station selection menu. Default thresholds are specific to each parameter.  

```{r}
column(12, 
       column(4, 
              selectInput('varsel', 'Select constituent:', choices = sort(c(tops, nutrs)), selected = 'ENT')
       ),
       column(4, 
              selectInput('tmdsel2', 'Filter by TMDL receiving waterbodies?', choices = c(F, T))
       ),
       column(4,
              renderUI({
                
                # inputs
                tmdsel2 <- as.logical(input$tmdsel2)
                varsel <- input$varsel
                
                if(!tmdsel2)
                  out <- selectInput('locsel2', 'Select watershed:', choices = unique(dwdat$Watershed))
                
                if(tmdsel2){
                  
                   # get receiving tmdl waterbodies 
                  grp <- tmdl %>% 
                    filter(Parameter %in% varsel) %>% 
                    pull(grp)
                  locs <- tmdldat %>% 
                    filter(Parameter %in% !!grp) %>% 
                    pull(Receiving) %>% 
                    unique
                  
                if(length(locs) == 0)
                  out <- 'No TMDLs for selection'
                
                if(length(locs) > 0)
                  out <- selectInput("locsel2", label = 'Select TMDL receiving waterbodies:', choices = locs)      
                
                }
                
                return(out)

              })
       ), 
)
column(12,
       column(4,  
              renderUI({
      
                # input
                spaflt <- wshpardat()
                stas <- unique(wshpardat()$StationCode)
                
                pickerInput(inputId = "stasel", label = 'Select stations:', choices = stas,
                  options = list(`actions-box` = TRUE, size = 20), selected = stas, multiple = TRUE)      
                
              })
       ),
       column(4, 
              selectInput('logsel', 'Log scale?', choices = c(F, T))
       ), 
       column(4,
              renderUI({
                
                # input
                hotrng3 <- hotrng3()
                
                numericInput('thrsel', 'Input threshold:', min = hotrng3[1], max = hotrng3[3], value = hotrng3[2])
                
              })
       )
)
column(12,
       column(4, 
              renderText({
         
                # input
                hotrng3 <- hotrng3()
                varthr2 <- varthr2()
                
                txt <- paste0("min ", hotrng3[1], ', max ', hotrng3[3])
     
                if(length(varthr2) == 0)
                  out <- paste0(txt, ', default selection is median (no threshold)')
                
                if(length(varthr2) != 0)
                  out <- paste0(txt, ', default selection is defined threshold')
                  
                return(out)
                       
              })
       )
)
```

### Observed

`r renderText({thrtx()})`

```{r}
output$obsp1 <- renderPlotly({obsp1()})
output$obsp2 <- renderPlotly({obsp2()})
output$obsp3 <- renderPlotly({obsp3()})
plotlyOutput('obsp1', height = "400px", width = "900px")
plotlyOutput('obsp2', height = "400px", width = "900px")
plotlyOutput('obsp3', height = "400px", width = "900px")
```
