---
output: 
  html_document:
      css: styles.css
runtime: shiny
---
  
# OC MS4 monitoring program, dry weather monitoring {.tabset}
  
```{r setup, message = F, warning = F, results = 'hide', echo = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, echo = F, fig.path = 'figs/', dev.args = list(family = 'serif'))

library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(lubridate)
library(leafsync)
library(viridisLite)
library(lubridate)
library(gridExtra)
library(stargazer)
library(EnvStats)
library(shiny)
library(kableExtra)
library(mgcv)
library(metR)
library(shinyWidgets)
library(plotly)
library(english)
library(vegan)
library(ggord)
library(patchwork)
library(plotly)
library(yaml)
library(pracma)
library(shinythemes)
library(shinycssloaders)
library(pwr)
library(purrr)

mptyps <- c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "Esri.WorldImagery", "OpenTopoMap")
# nutrients, top parameters (top ten plus those with thresholds)
nutrs <- c('Ammonia', 'Nitrate, Nitrite', 'Total Kjeldahl Nitrogen', 'Total Nitrogen', 'Orthophosphate', 'Total Phosphorus')

mapviewOptions(leafletHeight = 300)

prj <- 4326 # wgs84

source('R/funcs.R')

##




# ggplot themes

#dataframe needed to upload


#dwdat<- read_csv("upload.csv")


#static dataframes
data(thrsdat)
data(tmdldat)
data(dwdat)
  
# test<-dwdat%>%
#   slice(1:3000)
# 
# write.csv(test, "upload.csv")




# xy locs
# locs <- dwdat %>%
#   select(StationCode, Longitude, Latitude) %>%
#   unique
# 

# # date ranges
dts <- dwdat %>%
  pull(Date) %>%
  range

# obs ranges per station
obsrng <- dwdat %>%
  group_by(StationCode, Parameter) %>%
  summarise(n = n()) %>%
  pull(n) %>%
  range

# obs ranges per station, averages across parameters
obsave <- dwdat %>%
  group_by(StationCode, Parameter) %>%
  summarise(n = n()) %>%
  group_by(StationCode) %>%
  summarise(n = mean(n, na.rm = T)) %>%
  mutate(n = round(n, 1))
obsaverng <- obsave %>%
  pull(n) %>%
  range



# color palette for hotspot exceedances, as prop
hotcol <- colorNumeric(
  palette = rev(RColorBrewer::brewer.pal(11, 'RdYlBu')),
  na.color = 'yellow',
    domain = c(0, 100)
  )



tops <- table(dwdat$Parameter) %>% sort %>% rev %>% .[1:10] %>% names %>% sort
tops <- thrsdat %>% 
  pull(Parameter) %>% 
  c(., tops) %>% 
  unique
tops <- tops[!tops %in% nutrs]

# lookup list for parameters with thresholds and TMDL waterbodies
tmdl <- list(
  EC = 'EC', 
  Metals = c('Ag', 'As', 'Cd', 'Cr', 'Cu', 'Hg', 'Ni', 'Pb', 'Se', 'Zn'), 
  Nutrients = c('Ammonia', 'Nitrate, Nitrite'), 
  Pathogens = c('ENT', 'FC', 'TC'),
  Pesticides = c('Chlorpyrifos', 'Diazinon', 'Dimethoate', 'Fensulfothion', 'Malathion'), 
  TSS = 'TSS'
  ) %>% 
  enframe('grp', 'Parameter') %>% 
  unnest(Parameter)
```

```{r reactives}

  data <- reactive({
    req(input$file_upload)
    read.csv(input$file_upload$datapath)
    
  })

locs<-reactive({
  req(input$file_upload)
  
  
  locs <- data() 
  
  
  locs%>%
  select(StationCode, Longitude, Latitude) %>%
  unique
  
  locs<-as.data.frame(locs)
  
  
})


# get threshold variable
varthr <- reactive({
  
  # input
  varsel3 <- input$varsel3
  
  out <- thrsdat %>% 
    filter(Parameter %in% varsel3) %>% 
    pull(Threshold)
  
  return(out)
  
})

# hotspot input ranges
hotrng <- reactive({
  
  # inputs
  varsel3 <- input$varsel3
  varthr <- varthr()
  
  # values for slider
  qnts <- c(0, 0.5, 1)
  out <- dwdat %>% 
    filter(Parameter %in% varsel3) %>%
    pull(Result) %>% 
    quantile(probs = qnts, na.rm = T) %>%
    round(2)

  # make the default selection the threshold if applicable
  if(length(varthr) != 0)
    out[2] <- varthr
    
  return(out)
  
})



# hotspot data
hotdat <- reactive({
  
  
  # inputs
  cntsel <- input$cntsel
  varsel3 <- input$varsel3
  dtssel <- input$dtssel
  thrsel2 <- input$thrsel2
  tmdsel <- as.logical(input$tmdsel)
  locsel <- input$locsel
  req(thrsel2)

  # filter by tmdl waterbodies
  if(tmdsel 
     
     & length(locsel) 
     
     > 0){
    
    grp <- tmdl %>% 
      filter(Parameter %in% varsel3) %>% 
      pull(grp) 

    stas <- tmdldat %>% 
      filter(Parameter %in% !!grp) %>% 
      filter(Receiving %in% locsel) %>% 
      select(StationCode, Receiving)

    out <- dwdat %>% 
      inner_join(stas, by = 'StationCode') %>% 
      select(-Watershed) %>% 
      rename(Watershed = Receiving)
  
  }
  
  if(!tmdsel)
    out <- dwdat 
  
  req(exists('out'))
  
  # summary for all dates, stations by selected constituent
  out <- out %>% 
    filter(Parameter %in% varsel3) %>%
    group_by(Watershed, StationCode) %>% 
    mutate(n = n()) %>% 
    filter(Date >= dtssel[1] & Date <= dtssel[2]) %>%
    mutate(
      abv = case_when(
        Result > thrsel2 ~ 1, 
        T ~ 0
      )
    ) %>% 
    summarise(
      exceeds = sum(abv),
      n = unique(n)
      ) %>% 
    ungroup() %>% 
    mutate(
      exceeds = round(100 * exceeds / n, 0), 
      cols = hotcol(exceeds), 
      cexs = scales::rescale(exceeds, to = c(4, 17), from = c(0, 100))
    ) %>% 
    select(Watershed, StationCode, exceeds, n, cols, cexs) %>% 
    filter(n >= cntsel[1] & n <= cntsel[2]) %>% 
    unique
  
  return(out)
  
})

# map of exceedances
hotmap <- reactive({
  

  
 #locs<-locs()
  
  # input
  hotdat <- hotdat()
  locs<-locs()
  
  req(nrow(hotdat) > 0)
  
  tomap <- hotdat %>%
    left_join(locs, by = 'StationCode', copy=TRUE ) %>%
    st_as_sf(coords = c("Longitude", "Latitude"), crs = prj)

  # hover pt labels
  labs <- paste(tomap$StationCode, ': ', tomap$exceeds, ' % exceeding, ', tomap$n, ' total obs.')
  leglab <- "% exceeding"
  
  mapviewOptions(leafletHeight = 400)

  m <- mapview(tomap, cex = tomap$cexs, label = labs, legend = F, layer.name = F, col.regions = tomap$cols, homebutton = F, map.types = mptyps)
  
  # add legend
  out <- m@map %>% 
    addLegend("bottomright", pal = hotcol, title = leglab, opacity = 1, values = tomap$exceeds) 
  
  return(out)
  
})


```

This website presents an evaluation of the Orange County dry weather (ID/IC) dataset. The content is separated into three main tabs.  

* __Inventory__: Map-based and tabular summaries of monitoring effort and basic characteristics of the data
* __Hotspots__: Assessment of hotspot sites based on threshold exceedances over time
* __Analyses by waterbody__: A simple analysis of threshold exceedances for select sites shown for complete time series

Each main tab includes sub-tabs or drop-down menus for selecting and viewing different content.  Selected constituents are also available for each analysis.  Because the constituent list was large for some stations, a pre-selected list was created that included the top ten measured constituents for the entire data set and those those with defined thresholds.   Nutrients were also added to the list of constituents available for analysis.  The following shows which constituents were included in the website. 

* __Top ten, including those with thresholds__: `r paste(tops, collapse = ', ')`
* __Nutrients__: `r paste(nutrs, collapse = ', ')`

## Hotspots {.tabset .tabset-pills}

This tab show hotspots for stations that have many exceedances over the period of record.  Hotspots are defined by the number of instances when the observed data were above the threshold for the constituent, divided by the total number of observations at a site.  Sites with larger points colored in red have many exceedances, as compared to sites with small points in blue that have few exceedances.  Thresholds are defined targets for each constituent and were obtained from Orange County.  Median values are applied in the absence of a threshold, which applied only to TKN, TP, and TN.

### Individual constituents

```{r}




column(12, 
       column(12, 
              sidebarPanel(
        textOutput("Data control elements"),
        # mod_data_ui("data_selection"),
        fileInput("file_upload", "Import csv data", accept = ".csv")
        #),
         # Dataset filters
        # choice_ui("choice_ui_watershed", "Watershed"),
        # choice_ui("choice_ui_station", "Station"),
        # choice_ui("choice_ui_parameter", "Parameter"),
        # choice_ui("choice_ui_unit", "Unit"),
        # choice_ui("choice_ui_fraction", "Fraction")
        ),
        
        ))


column(12, 
       column(4, 
              selectInput('varsel3', 'Select constituent:', choices = sort(c(tops, nutrs)), selected = 'ENT')
       ),
       column(8,
              sliderInput('dtssel', 'Select date range:', min = dts[1], max = dts[2], value = dts, timeFormat="%Y-%m", width = '600px')
       )
)
column(12,
       column(4,
              selectInput('tmdsel', 'Filter by TMDL receiving waterbodies?', choices = c(F, T))
       ),
       column(8,
              sliderInput('cntsel', 'Filter stations by total n:', min = 0, max = obsrng[2], value = c(0, obsrng[2]), width = '600px')
       )
)
column(12,
       column(4,
              renderUI({


            #locs<-locs()
                # input
                varsel3 <- input$varsel3
                tmdsel <- as.logical(input$tmdsel)

                req(tmdsel)

                # get receiving tmdl waterbodies
                grp <- tmdl %>%
                  filter(Parameter %in% varsel3) %>%
                  pull(grp)
                locs <- tmdldat %>%
                  filter(Parameter %in% !!grp) %>%
                  pull(Receiving) %>%
                  unique

                if(length(locs) == 0)
                  out <- 'No TMDLs for selection'

                if(length(locs1) > 0)
                  out <- pickerInput(inputId = "locsel", label = 'Select TMDL receiving waterbodies:', choices = locs,
                    options = list(`actions-box` = TRUE, size = 20), selected = locs, multiple = TRUE)

                return(out)

              })
       ),
       column(4,
              renderUI({
                
                # input
                hotrng <- hotrng()
                
                numericInput('thrsel2', 'Input threshold:', min = hotrng[1], max = hotrng[3], value = hotrng[2])
                
              })
       ), 
       column(4, 
              renderText({
         
                # input
                hotrng <- hotrng()
                varthr <- varthr()
                
                txt <- paste0("min ", hotrng[1], ', max ', hotrng[3])
     
                if(length(varthr) == 0)
                  out <- paste0(txt, ', default selection is median (no threshold)')
                
                if(length(varthr) != 0)
                  out <- paste0(txt, ', default selection is defined threshold')
                  
                return(out)
                       
              })
       )
)
```

This map shows the percentage of observations above a threshold for `r renderText({input$varsel3})` for the period of record between `r renderText(as.character(input$dtssel[1]))` and `r renderText(as.character(input$dtssel[2]))`.  Sites are also filtered based on the number of total observations, where sites sites with as few as `r renderText(input$cntsel[1])` and as many as `r renderText(input$cntsel[2])` observations are shown below.  The threshold defining exceedances for `r renderText({input$varsel3})` is shown above.  

Sites can also be filtered based on the receiving waterbody with a TMDL.  Selecting "TRUE" from the drop-down list will create an additional list to filter the stations by the receiving waterbody. By default, all stations are shown.

```{r}


renderLeaflet(hotmap())
```
