---
output: 
  html_document:
      css: styles.css
runtime: shiny
---
  
# App 4, Constituent Hotspots {.tabset}
  
```{r setup, message = F, warning = F, results = 'hide', echo = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, echo = F, fig.path = 'figs/', dev.args = list(family = 'serif'))

library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(lubridate)
library(leafsync)
library(viridisLite)
library(lubridate)
library(gridExtra)
library(stargazer)
library(EnvStats)
library(shiny)
library(kableExtra)
library(mgcv)
library(metR)
library(shinyWidgets)
library(plotly)
library(english)
library(vegan)
library(ggord)
library(patchwork)
library(plotly)
library(yaml)
library(pracma)
library(shinythemes)
library(shinycssloaders)
library(pwr)
library(purrr)
library(DT)
library(writexl)

mptyps <- c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "Esri.WorldImagery", "OpenTopoMap")
# nutrients, top parameters (top ten plus those with thresholds)
nutrs <- c('Ammonia', 'Nitrate, Nitrite', 'Total Kjeldahl Nitrogen', 'Total Nitrogen', 'Orthophosphate', 'Total Phosphorus')

mapviewOptions(leafletHeight = 300)

prj <- 4326 # wgs84

source('R/funcs.R')

##




# ggplot themes

#dataframe needed to upload


#dwdat<- read_csv("upload.csv")


#static dataframes
data(thrsdat)
data(tmdldat)
data(dwdat)
  
# test<-dwdat%>%
#   slice(1:3000)
# 
# write.csv(test, "upload.csv")




# xy locs
# locs <- dwdat %>%
#   select(StationCode, Longitude, Latitude) %>%
#   unique
# 

# date ranges
# dts <- dwdat %>%
#   pull(Date) %>%
#   range

# # obs ranges per station
# obsrng <- dwdat %>%
#   group_by(StationCode, Parameter) %>%
#   summarise(n = n()) %>%
#   pull(n) %>%
#   range
# 
# # obs ranges per station, averages across parameters
# obsave <- dwdat %>%
#   group_by(StationCode, Parameter) %>%
#   summarise(n = n()) %>%
#   group_by(StationCode) %>%
#   summarise(n = mean(n, na.rm = T)) %>%
#   mutate(n = round(n, 1))
# obsaverng <- obsave %>%
#   pull(n) %>%
#   range
# 
# 

# color palette for hotspot exceedances, as prop
hotcol <- colorNumeric(
  palette = rev(RColorBrewer::brewer.pal(11, 'RdYlBu')),
  na.color = 'yellow',
    domain = c(0, 100)
  )



# tops <- table(dwdat$Parameter) %>% sort %>% rev %>% .[1:10] %>% names %>% sort
# 
# 
# tops <- thrsdat %>%
#   pull(Parameter) %>%
#   c(., tops) %>%
#   unique
# tops <- tops[!tops %in% nutrs]

# lookup list for parameters with thresholds and TMDL waterbodies
tmdl <- list(
  EC = 'EC', 
  Metals = c('Ag', 'As', 'Cd', 'Cr', 'Cu', 'Hg', 'Ni', 'Pb', 'Se', 'Zn'), 
  Nutrients = c('Ammonia', 'Nitrate, Nitrite'), 
  Pathogens = c('ENT', 'FC', 'TC'),
  Pesticides = c('Chlorpyrifos', 'Diazinon', 'Dimethoate', 'Fensulfothion', 'Malathion'), 
  TSS = 'TSS'
  ) %>% 
  enframe('grp', 'Parameter') %>% 
  unnest(Parameter)
```

```{r reactives}


#Template 

StationCode <- c("ANAB0151@WES")
Watershed <- c("Santa Ana River")
Date<-c("2016-05-23")
Parameter<-c("Cu")
Result<-c("23")
Units<-c("ug/L")
Longitude<-c("Longitude")
Latitude<-c("Latitude")


df <- data.frame(StationCode,Watershed,Date,Parameter,Result,Units,Longitude,Latitude)



output$dl <- downloadHandler(
    filename = function() { "Template.xlsx"},
    content = function(file) {write_xlsx(df, path = file)}
  )



  data <- reactive({
    req(input$file_upload)
    read.csv(input$file_upload$datapath)
    
  })

filtered_dataframe<-reactive({
  
   req(input$file_upload)
   data<-read.csv(input$file_upload$datapath)
  
    #quantile_c<-input$quantile_check
    parameter_c<-input$varsel3
    #watershed_c<-input$Watershed_check
    
    
  data%>%
    #filter(StationCode %in% quantile_c)%>%
    filter(Parameter %in% parameter_c)#%>%
    #filter(Watershed %in% watershed_c)
  
  
})


# tops <- reactive({
#   
#   req(input$file_upload)
#    data<-read.csv(input$file_upload$datapath)
#   
#   tops<-table(data$Parameter) %>% sort %>% rev %>% .[1:10] %>% names %>% sort
#   
#   tops <- thrsdat %>% 
#   pull(Parameter) %>% 
#   c(., tops) %>% 
#   unique
# tops <- tops[!tops %in% nutrs]
# 
# return(tops)
#   
# })


  
  

locs<-reactive({
  req(input$file_upload)
  
  
  locs <- filtered_dataframe()
  
  
  locs%>%
  select(StationCode, Longitude, Latitude) %>%
  unique
  
  print(locs)
  
return(locs)
  


  
})



  
  
  
  

output$my_table<-renderDT({
  
  data()
 
  
})

# dts<-reactive({
#   req(input$file_upload)
# 
# 
#   dts <- data()
# 
# 
#  dts %>%
#   pull(Date) %>%
#   range
# 
#   print(dts)
# 
# return(dts)
# 
# 
# })


# get threshold variable
varthr <- reactive({
  
  # input
  varsel3 <- input$varsel3
  
  out <- thrsdat %>% 
    filter(Parameter %in% varsel3) %>% 
    pull(Threshold)
  
  return(out)
  
})

# hotspot input ranges
hotrng <- reactive({
  
  req(input$file_upload)
  
  data<-data()
  
  # inputs
  varsel3 <- input$varsel3
  varthr <- varthr()
  
  # values for slider
  qnts <- c(0, 0.5, 1)
  out <- data %>% 
    filter(Parameter %in% varsel3) %>%
    pull(Result) %>% 
    quantile(probs = qnts, na.rm = T) %>%
    round(2)

  # make the default selection the threshold if applicable
  if(length(varthr) != 0)
    out[2] <- varthr
    
  return(out)
  
})



# hotspot data
hotdat <- reactive({
  
  
  data<-data()
  
  # inputs
  #cntsel <- input$cntsel
  varsel3 <- input$varsel3
  #dtssel <- input$dtssel
  thrsel2 <- input$thrsel2
  tmdsel <- as.logical(input$tmdsel)
  locsel <- input$locsel
  req(thrsel2)

  # filter by tmdl waterbodies
  if(tmdsel 
     
     & length(locsel) 
     
     > 0){
    
    grp <- tmdl %>% 
      filter(Parameter %in% varsel3) %>% 
      pull(grp) 

    stas <- tmdldat %>% 
      filter(Parameter %in% !!grp) %>% 
      filter(Receiving %in% locsel) %>% 
      select(StationCode, Receiving)

    out <- data %>% 
      inner_join(stas, by = 'StationCode') %>% 
      select(-Watershed) %>% 
      rename(Watershed = Receiving)
  
  }
  
  if(!tmdsel)
    out <- data 
  
  req(exists('out'))
  
  # summary for all dates, stations by selected constituent
  out <- out %>% 
    filter(Parameter %in% varsel3) %>%
    group_by(Watershed, StationCode) %>% 
    mutate(n = n()) %>% 
    #filter(Date >= dtssel[1] & Date <= dtssel[2]) %>%
    mutate(
      abv = case_when(
        Result > thrsel2 ~ 1, 
        T ~ 0
      )
    ) %>% 
    summarise(
      exceeds = sum(abv),
      n = unique(n)
      ) %>% 
    ungroup() %>% 
    mutate(
      exceeds = round(100 * exceeds / n, 0), 
      cols = hotcol(exceeds), 
      cexs = scales::rescale(exceeds, to = c(4, 17), from = c(0, 100))
    ) %>% 
    select(Watershed, StationCode, exceeds, n, cols, cexs) %>% 
    #filter(n >= cntsel[1] & n <= cntsel[2]) %>% 
    unique
  
  return(out)
  
  
})

# map of exceedances
hotmap <- eventReactive(list(input$go_quality),{
  

  
 #locs<-locs()
  
  # input
  hotdat <- hotdat()
  locs<-locs()
  
  req(nrow(hotdat) > 0)
  
  tomap <- hotdat %>%
    inner_join(locs, by = 'StationCode', copy=TRUE )
  print("tomap")
  print(tomap)
  tomap <- tomap %>%
    st_as_sf(coords = c("Longitude", "Latitude"), crs = prj)
  


  # hover pt labels
  labs <- paste(tomap$StationCode, ':', tomap$exceeds, ' % exceeding, ', tomap$n, ' total obs.')
  leglab <- "% exceeding"
  
  mapviewOptions(leafletHeight = 600,leafletWidth = 800)

  m <- mapview(tomap, cex = tomap$cexs, label = labs, legend = F, layer.name = F, col.regions = tomap$cols, homebutton = F, map.types = mptyps)
  
  # add legend
  out <- m@map %>% 
    addLegend("bottomright", pal = hotcol, title = leglab, opacity = 1, values = tomap$exceeds) 
  
  return(out)
  
})


```


This tab show hotspots for each constituent in a given station and watershed. Hotspots are defined by the number of instances when the observed data were above the threshold for the constituent, divided by the total number of observations at a site.  Sites with larger points colored in red have many exceedances, as compared to sites with small points in blue that have few excedances.  Thresholds are defined targets for each constituent and were obtained from the international BMP database.  Median values are applied in the absence of a threshold, which applied only to TKN, TP, and TN.

### Individual constituents

```{r}




column(12, 
       column(10, 
              sidebarPanel(width=10,
        downloadButton("dl", "Download Template Here"),
        textOutput("Data control elements"),
        # mod_data_ui("data_selection"),
        fileInput("file_upload", "Import csv data", accept = ".csv"),
        renderUI({  
           
           data<-data()
           data<-data%>%
             mutate(StationCode = factor(StationCode))%>%
             mutate(Parameter = factor(Parameter))%>%
             mutate(Watershed = factor(Watershed))
             #print(data)
           
  
           # pickerInput(inputId = "Watershed_check", 
           #             label = "Watershed (Select up tp 3):",
           #             choices = levels(data$Watershed),
           #             selected = levels(data$Watershed),
           #             options = list(`actions-box` = TRUE), 
           #             multiple = TRUE)
           
            pickerInput(inputId = "varsel3", 
                       label = "Constituent:",
                       choices = levels(data$Parameter),
                       selected = levels(data$Parameter),
                       options = list(`actions-box` = TRUE), 
                       multiple = FALSE)
            
           
         }),
        #selectInput('varsel3', 'Select constituent:', choices = sort(c(tops, nutrs)), selected = 'ENT'),
        selectInput('tmdsel', 'Filter by TMDL receiving waterbodies?', choices = c(F, T)),
        renderUI({


            #locs<-locs()
                # input
                varsel3 <- input$varsel3
                tmdsel <- as.logical(input$tmdsel)

                req(tmdsel)

                # get receiving tmdl waterbodies
                grp <- tmdl %>%
                  filter(Parameter %in% varsel3) %>%
                  pull(grp)
                locs <- tmdldat %>%
                  filter(Parameter %in% !!grp) %>%
                  pull(Receiving) %>%
                  unique

                if(length(locs) == 0)
                  out <- 'No TMDLs for selection'

                if(length(locs) > 0)
                  out <- pickerInput(inputId = "locsel", label = 'Select TMDL receiving waterbodies:', choices = locs,
                    options = list(`actions-box` = TRUE, size = 20), selected = locs, multiple = TRUE)

                return(out)

              }),
        
        
        renderUI({
                
                # input
                hotrng <- hotrng()
                
                numericInput('thrsel2', 'Input threshold:', min = hotrng[1], max = hotrng[3], value = hotrng[2])
                
              }),
        
        actionButton("go_quality", "Plot Current Selection"),
         # Dataset filters
        # choice_ui("choice_ui_watershed", "Watershed"),
        # choice_ui("choice_ui_station", "Station"),
        # choice_ui("choice_ui_parameter", "Parameter"),
        # choice_ui("choice_ui_unit", "Unit"),
        # choice_ui("choice_ui_fraction", "Fraction")
        )),
)

column(12,
       column(12,leafletOutput('mymap', width=800))
       
       
       )


 # column(12, 
 #        column(6, 
 #                renderLeaflet(hotmap()),
 #        ),
        #column(6,
               #dataTableOutput("my_table")
#        # sliderInput('dtssel', 'Select date range:', min = dts[1], max = dts[2], value = dts, timeFormat="%Y-%m", width = '600px')
        #)
 #)
# column(12,
#        column(4,
#               selectInput('tmdsel', 'Filter by TMDL receiving waterbodies?', choices = c(F, T))
#        ),
       #column(8,
              #sliderInput('cntsel', 'Filter stations by total n:', min = 0, max = obsrng[2], value = c(0, obsrng[2]), width = '600px')
       #)
#)
column(12,
       # column(4,
       #        renderUI({
       # 
       # 
       #      #locs<-locs()
       #          # input
       #          varsel3 <- input$varsel3
       #          tmdsel <- as.logical(input$tmdsel)
       # 
       #          req(tmdsel)
       # 
       #          # get receiving tmdl waterbodies
       #          grp <- tmdl %>%
       #            filter(Parameter %in% varsel3) %>%
       #            pull(grp)
       #          locs <- tmdldat %>%
       #            filter(Parameter %in% !!grp) %>%
       #            pull(Receiving) %>%
       #            unique
       # 
       #          if(length(locs) == 0)
       #            out <- 'No TMDLs for selection'
       # 
       #          if(length(locs) > 0)
       #            out <- pickerInput(inputId = "locsel", label = 'Select TMDL receiving waterbodies:', choices = locs,
       #              options = list(`actions-box` = TRUE, size = 20), selected = locs, multiple = TRUE)
       # 
       #          return(out)
       # 
       #        })
       # ),
       # column(4,
       #        renderUI({
       #          
       #          # input
       #          hotrng <- hotrng()
       #          
       #          numericInput('thrsel2', 'Input threshold:', min = hotrng[1], max = hotrng[3], value = hotrng[2])
       #          
       #        })
       # ), 
       column(4, 
              renderText({
         
                # input
                hotrng <- hotrng()
                varthr <- varthr()
                
                txt <- paste0("min ", hotrng[1], ', max ', hotrng[3])
     
                if(length(varthr) == 0)
                  out <- paste0(txt, ', default selection is median (no threshold)')
                
                if(length(varthr) != 0)
                  out <- paste0(txt, ', default selection is defined threshold')
                  
                return(out)
                       
              })
       )
)
```



```{r}


output$mymap<-renderLeaflet(hotmap())
```
