---
output: 
    html_document:
        css: styles.css
runtime: shiny
---

# OC MS4 monitoring program, mass emissions loading {.tabset}

```{r setup, message = F, warning = F, results = 'hide', echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, echo = F, fig.path = 'figs/', dev.args = list(family = 'serif'))

library(tidyverse)
library(sf)
library(mapview)
library(lubridate)
library(leaflet)
library(leafsync)
library(viridisLite)
library(lubridate)
library(gridExtra)
library(stargazer)
library(EnvStats)
library(shiny)
library(kableExtra)
library(mgcv)
library(metR)
library(vegan)
library(ggord)
library(patchwork)
library(ggrepel)
library(ggExtra)
library(plotly)
library(matrixStats)

mptyps <- c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "Esri.WorldImagery", "OpenTopoMap")

mapviewOptions(leafletHeight = 300)

prj <- 4326 # wgs84

source('R/funcs.R')

##
# ggplot themes

thm1 <- theme_bw(base_size = 16) + 
  theme(
    strip.background = element_blank(), 
    strip.placement = 'outside', 
    strip.text = element_text(hjust = 0),
    axis.title.x = element_blank(), 
    legend.title = element_blank(), 
    legend.position = 'bottom', 
    panel.grid = element_blank()
  )

thm2 <- theme_bw(base_size = 16) + 
  theme(
    strip.background = element_blank(), 
    strip.placement = 'outside', 
    legend.position = 'bottom', 
    panel.grid = element_blank()
  )

thm3 <- theme_bw(base_size = 12) + 
  theme(
    strip.background = element_blank(), 
    strip.placement = 'outside', 
    legend.position = 'bottom', 
    panel.grid = element_blank()
  )

pbase <- theme_bw(base_family = 'serif') +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
    axis.text.y = element_text(size = 10),
    legend.position = 'top',
    legend.direction = 'horizontal',
    # plot.margin = unit(c(4,4,0,0), "lines"),
    strip.background = element_blank(), 
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
    panel.background = element_rect(fill = 'black')
  ) 

data(lddat)

# constituents
metals <- c("Ag", "As", "Cd", "Cr", "Cu", "Fe", "Hg", "Ni", "Pb", "Se", 
            "Zn")
nutrs <- c('Ammonia', 'Nitrate, Nitrite', 'Total Kjeldahl Nitrogen', 'Orthophosphate', 'Total Phosphorus')
```

```{r reactives}
# inventory table
invtab <- reactive({
  
  # input
  consel1 <- input$consel1
  tabtyp <- input$tabtyp

  # filter by parameters
  tofilt <- eval(parse(text = consel1))
  sums <- lddat %>% 
    filter(Parameter %in% tofilt)
  
  if(tabtyp == 'Number of observations'){
    
    sums <- sums %>% 
      dplyr::count(Parameter, StationCode) %>% 
      bind_rows(group_by(., Parameter) %>%
                summarise(n=sum(n)) %>%
                mutate(StationCode='Total')) %>%
      bind_rows(group_by(., StationCode) %>%
                summarise(n=sum(n)) %>%
                mutate(Parameter='Total')) %>%
      spread(Parameter, n, fill=0) %>% 
      select_at(c('StationCode', tofilt, 'Total'))
  
    alltot <- sums %>% 
      filter(StationCode != 'Total')
    coltot <- sums %>% 
      filter(StationCode == 'Total')
    
    totab <- bind_rows(alltot, coltot)
    
    cap <- 'Number of observations at each site'
  
  }
        
  if(tabtyp == 'Variance'){
    
    totab <- sums %>% 
      group_by(Parameter, StationCode) %>% 
      summarise(vr = round(var(Result, na.rm = T), 2)) %>% 
      spread(Parameter, vr, fill = NA)
    
    cap <- 'Variance at each site'
    
  }
  
  if(tabtyp == 'Median absolute deviation'){
    
    totab <- sums %>% 
      group_by(Parameter, StationCode) %>% 
      summarise(md = round(mad(Result, na.rm = T), 2)) %>% 
      spread(Parameter, md, fill = NA)
    
    cap <- 'Median absolute deviation at each site'
    
  }
  
  HTML(knitr::kable(totab, format = 'html', caption = cap) %>% 
      kable_styling(full_width = T, font_size = 14))  
    
})

# selected station, parameter data
stapardat <- reactive({
  
  # inputs
  stasel <- input$stasel
  varsel <- input$varsel

  req(varsel)
  req(stasel)
  
  out <- lddat %>% 
    filter(StationCode %in% stasel) %>% 
    filter(Parameter %in% varsel) %>% 
    mutate(
      Year = year(Date), 
      Month = month(Date, label = T)
    ) 

  validate(
    need(nrow(out) > 0, 'No data in selection')
  )    
  
  return(out)
  
})

# first observed plot
obsp1 <- reactive({
  
  # inputs
  stapardat <- stapardat()
  stasel <- input$stasel

  toplo <- stapardat
  ttl <- unique(toplo$Parameter)
  
  # load
  p1 <- ggplot(toplo, aes(x = Date, y = Result)) + 
    geom_line() + 
    geom_point() + 
    thm1 + 
    # scale_y_log10() + 
    labs(
      subtitle = stasel, 
      title = ttl, 
      y = 'Load'
    )
  
  # flow
  p2 <- ggplot(toplo, aes(x = Date, y = Flow)) + 
    geom_line() + 
    geom_point() + 
    thm1 + 
    # scale_y_log10() + 
    labs(
      y = 'Flow volume'
    )
  
  out <- p1 + p2 + plot_layout(ncol = 1)
  
  return(out)
  
})

# second observed plot
obsp2 <- reactive({
  
  # inputs
  stapardat <- stapardat()
  stasel <- input$stasel
  
  toplo <- stapardat
  ttl <- unique(toplo$Parameter)

  p <- ggplot(toplo, aes(x = factor(Year), y = Result, group = Year)) + 
    geom_boxplot() + 
    thm1 +
    # scale_y_log10() + 
    # scale_colour_manual(values = c('tomato1', 'black')) + 
    labs(
      subtitle = stasel,
      title = paste(ttl, 'by year'), 
      y = 'Load'
    )

  return(p)
  
})

# third observed plot
obsp3 <- reactive({
  
  # inputs
  stapardat <- stapardat()
  stasel <- input$stasel
  
  toplo <- stapardat
  ttl <- unique(toplo$Parameter)

  p <- ggplot(toplo, aes(x = Month, y = Result, group = Month)) + 
    geom_boxplot() + 
    thm1 +
    # scale_y_log10() + 
    # scale_colour_manual(values = c('tomato1', 'black')) + 
    labs(
      subtitle = stasel, 
      title = paste(ttl, 'by month'), 
      y = 'Load'
    )

  return(p)
  
})

# trend data
trnddat <- reactive({
  
  # input
  stapardat <- stapardat()
  
  out <- try({stapardat %>%
    group_by(Parameter, Year) %>%
    summarise(
      Result = mean(Result, na.rm = T),
    ) %>%
    group_by(Parameter) %>%
    mutate(
      avg = mean(Result, na.rm = T),
      dev = Result - avg
    ) %>%
    ungroup %>%
    mutate(
      avg = round(avg, 2),
      Parameter = paste0(Parameter, ' (', avg, ')')
    )
  })
  
  return(out)
  
})

# first trend plot
trndp1 <- reactive({
  
  # input
  trnddat <- trnddat()
  stasel <- input$stasel
  
  validate(
    need(!inherits(trnddat, 'try-error'), 'Insufficient data')
  )
  
  subttl <- paste0(stasel, ', average in parentheses on y-axis')
  
  p <- ggplot(trnddat, aes(x = Year, y = dev, fill = dev)) +
    geom_bar(stat = 'identity', colour = 'grey') +
    scale_fill_gradient2('Deviations from average (+/-)', low = 'tomato1', mid = 'grey90', high = 'lightgreen', midpoint = 0) +
    facet_wrap(~Parameter, strip.position = 'left', ncol = 1, scales = 'free_y') +
    thm1 +
    theme(axis.title.y = element_blank()) +
    geom_smooth(method = 'lm', se = F, linetype = 'dashed', color = 'black') +
    # scale_y_log10() +
    # scale_y_continuous('Density (fish/100ft)') +
    geom_hline(aes(yintercept = 0)) +
    labs(
      title = 'Deviation from annual averages by constituent (mg/L)',
      subtitle = subttl
    ) + 
    guides(fill = guide_colourbar(barwidth = 15, barheight = 0.5))
  
  return(p)
    
})

# first trend table
trndtab1 <- reactive({
  
  # input
  trnddat <- trnddat()
  
  validate(
    need(!inherits(trnddat, 'try-error'), 'Insufficient data')
  )
  
  mods <- trnddat %>%
    group_by(Parameter) %>%
    nest %>%
    mutate(
      mod = purrr::map(data, function(x) lm(dev ~ Year, x))
    ) %>%
    ungroup %>%
    mutate(
      Parameter = gsub('\\s\\(.*\\)$', '', Parameter)
    )
  
  out <- stargazer(mods$mod, type = 'html',
            covariate.labels = c('Year', 'Intercept'),
            column.labels = mods$Parameter,
            omit.stat = c('adj.rsq', 'ser'),
            column.sep.width = '10pt',
            dep.var.labels.include = F,
            digits = 2
            ) %>% 
    HTML()
  
  return(out)

})

# second trends table
trndtab2 <- reactive({
  
  # input
  stapardat <- stapardat()
  
  totabs1kn <- stapardat %>%
    group_by(Parameter, Year) %>%
    summarise(
      Result = mean(Result, na.rm = T),
    ) %>%
    group_by(Parameter) %>%
    mutate(
      avg = mean(Result, na.rm = T),
      dev = Result - avg
    ) %>%
    nest %>%
    mutate(
      res = purrr::map(data, function(x){
  
        knout <- kendallTrendTest(dev ~ Year, x)
        outest <- round(knout$estimate, 2)
        outpval <- p_ast(knout$p.value)
        nval <- nrow(na.omit(x))
        out <- c(n = nval, outest, pval = outpval) %>%
          data.frame %>%
          t %>%
          data.frame %>%
          select(-intercept)
        return(out)
  
      })
  
    ) %>%
    select(-data) %>%
    unnest(res)

  out <- HTML(knitr::kable(totabs1kn, format = 'html') %>%
    kable_styling(full_width = T, font_size = 14))
  
  return(out)
  
})

# comparison of time series for stations
stapardat2 <- reactive({
  
  # inputs
  stasel2 <- input$stasel2
  stasel3 <- input$stasel3
  varsel2 <- input$varsel2

  req(varsel2)
  
  out <- lddat %>% 
    filter(StationCode %in% c(stasel2, stasel3)) %>% 
    mutate(
      Year = year(Date), 
      Month = month(Date, label = T), 
      StationCode = factor(StationCode, levels = c(stasel2, stasel3))
    ) %>%
    filter(Parameter %in% varsel2)
    
  return(out)
  
})

# fourth observed plot
obsp4 <- reactive({
  
  # inputs
  stapardat2 <- stapardat2()
  
  toplo <- stapardat2
  ttl <- unique(toplo$Parameter)
  
  p <- ggplot(toplo, aes(x = Date, y = Result)) + 
    geom_line() + 
    facet_wrap(~StationCode, ncol = 1) + 
    thm1 + 
    # scale_y_log10() + 
    labs(
      title = ttl, 
      y = 'Load'
    )
  
  return(p)
  
})

# fifth observed plot
obsp5 <- reactive({
  
  # inputs
  stapardat2 <- stapardat2()
  
  toplo <- stapardat2
  ttl <- unique(toplo$Parameter)

  p <- ggplot(toplo, aes(x = factor(Year), y = Result, group = Year)) + 
    geom_boxplot() + 
    facet_wrap(~StationCode, ncol = 1) + 
    thm1 +
    # scale_y_log10() + 
    # scale_colour_manual(values = c('tomato1', 'black')) + 
    labs(
      title = paste(ttl, 'by year'), 
      y = 'Load'
    )

  return(p)
  
})

# sixth observed plot
obsp6 <- reactive({
  
  # inputs
  stapardat2 <- stapardat2()
  
  toplo <- stapardat2
  ttl <- unique(toplo$Parameter)

  p <- ggplot(toplo, aes(x = Month, y = Result, group = Month)) + 
    geom_boxplot() + 
    facet_wrap(~StationCode, ncol = 1) + 
    thm1 +
    # scale_y_log10() + 
    # scale_colour_manual(values = c('tomato1', 'black')) + 
    labs(
      title = paste(ttl, 'by month'), 
      y = 'Load'
    )

  return(p)
  
})

# dissimilarity data
disdat <- reactive({
  
  # input
  varsel2 <- input$varsel2
  
  req(varsel2)

  dissim <- ldprp %>% 
    filter(Parameter %in% varsel2) %>% 
    select(-Watershed, -Parameter) %>% 
    spread(StationCode, Result) %>% 
    column_to_rownames('Date') %>% 
    decostand(method = 'log', na.rm = T) %>% 
    decostand(method= 'standardize', na.rm = T) %>%
    rownames_to_column('Date') %>% 
    gather('StationCode', 'Result', -Date) %>% 
    spread(Date, Result) %>% 
    column_to_rownames('StationCode') %>% 
    vegdist(method = 'euclidean', na.rm = T) %>%
    as.matrix %>%
    as.data.frame %>%
    rownames_to_column('StationCode') %>%
    gather('StationCode2', 'dist', -StationCode) %>%
    left_join(siteshd, by = 'StationCode') %>%
    arrange(Watershed, dist) %>%
    mutate(
      Watershed = factor(Watershed, levels = unique(Watershed)),
      dist = ifelse(StationCode == StationCode2, NA, dist)
    )
  
  return(dissim)
  
})

# dissimilarity data, all constituents
discondat <- reactive({
   
  # input
  consel3 <- input$consel3

  sels <- eval(parse(text = consel3))
           
  dissim <- ldprp %>% 
    filter(Parameter %in% sels) %>% 
    select(-Watershed) %>% 
    group_by(Parameter) %>% 
    nest() %>% 
    mutate(
      dis = purrr::map(data, function(data){
        
        out <- data %>% 
          spread(StationCode, Result) %>% 
          column_to_rownames('Date') %>% 
          decostand(method = 'log', na.rm = T) %>% 
          decostand(method= 'standardize', na.rm = T) %>%
          rownames_to_column('Date') %>% 
          gather('StationCode', 'Result', -Date) %>% 
          spread(Date, Result) %>% 
          column_to_rownames('StationCode') %>% 
          vegdist(method = 'euclidean', na.rm = T) %>%
          as.matrix %>%
          as.data.frame %>%
          rownames_to_column('StationCode') %>%
          gather('StationCode2', 'dist', -StationCode) %>%
          left_join(siteshd, by = 'StationCode') %>%
          arrange(Watershed, dist) %>%
          mutate(
            Watershed = factor(Watershed, levels = unique(Watershed)),
            dist = ifelse(StationCode == StationCode2, NA, dist)
          )
        
        return(out)
        
      })
    ) %>% 
    select(-data) %>% 
    unnest(dis) %>% 
    na.omit %>% 
    group_by(Parameter, StationCode) %>% 
    summarise(dist = mean(dist, na.rm = T)) %>% 
    group_by(StationCode) %>% 
    summarise(dist = mean(dist, na.rm = T))
    
  return(dissim)
  
})

# multivariate comparison plots
mulplo <- reactive({
  
  # input
  disdat <- disdat()
  stasel2 <- input$stasel2
  stasel3 <- input$stasel3
  varsel2 <- input$varsel2
  consel3 <- input$consel3
  
  req(varsel2)
  
  ##
  # PCA
  sels <- eval(parse(text = consel3))
           
  toord <- ldprp %>% 
    filter(Parameter %in% !!sels) %>% 
    spread(Parameter, Result) %>% 
    na.omit %>%
    unite('rwnm', StationCode, Date) %>% 
    column_to_rownames('rwnm')
  wshd <- toord$Watershed
  toord <- toord %>% 
    select(-Watershed) %>% 
    decostand(method = 'log', na.rm = T) %>% 
    decostand(method= 'standardize', na.rm = T) %>% 
    select(which(!is.na(apply(., 2, var))))

  ord <- toord %>% 
    prcomp(scale. = T, center = T)

  # station subset
  vals <- ord$x %>% 
    as.data.frame %>% 
    rownames_to_column() %>% 
    mutate(StationCode = gsub('^(.*)\\_.*$', '\\1', rowname)) %>% 
    filter(StationCode %in% c(stasel2, stasel3)) %>% 
    rename(
      one = PC1, 
      two = PC2
    ) %>% 
    mutate(
      StationCode = factor(StationCode, levels = c(stasel2, stasel3))
    )

  p1 <- ggord(ord, size = 1, grp_in = wshd, alpha = 0.4, vec_ext = 5, coord_fix = F, parse = F) + 
    scale_fill_viridis_d(guide_legend(ncol = 1)) + 
    geom_point(data = vals, aes(alpha = StationCode), size = 3) +
    scale_alpha_manual(values = c(0.2, 0.8)) + 
    theme(
      legend.position = 'top',
      legend.title = element_blank(),
      legend.box = 'vertical'
      )
  
  ##
  # dissimilarity data and plots
  outplo <- disdat %>% 
    filter(StationCode %in% stasel2) %>% 
    filter(StationCode2 %in% stasel3)
  
  # dissimilarity ranges
  disrng <- range(disdat$dist, na.rm = T)
  
  dismid <- disrng[1] + (disrng[2] - disrng[1]) / 2
  subttl <- paste0("Dissimilarity between ", stasel2, " and ", stasel3, " ", round(outplo$dist, 1))
  
  p2 <- ggplot(disdat) +
    geom_tile(aes(x = StationCode, y = StationCode2, fill = dist), colour = 'black') +
    geom_tile(data = outplo, aes(x = StationCode, y = StationCode2), fill = NA, colour = 'black', size = 2) +
    scale_x_discrete('', expand = c(0, 0)) +
    scale_y_discrete('', expand = c(0, 0)) +
    scale_fill_gradient2('Dissimilarity between stations', low = 'lightblue', mid = 'white', high = 'tomato1', midpoint = dismid, limits = disrng) +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10, label.theme = element_text(size = 11, angle = 0))) +
    labs(
      title = varsel2, 
      subtitle = subttl
    ) +
    pbase

  p1 + p2 + plot_layout(ncol = 2)
      
})

# all trends across stations for selected constituent
alltrnd <- reactive({
  
  # input
  varsel3 <- input$varsel3
  
  req(varsel3)
  
  out <- lddat %>%
    filter(Parameter %in% varsel3) %>%
    mutate(
      Year = year(Date),
      Month = month(Date, label = T)
    ) %>%
    group_by(StationCode, Year) %>%
    summarise(
      Result = mean(Result, na.rm = T),
    ) %>%
    group_by(StationCode) %>%
    # filter(n() == 10) %>%
    mutate(
      avg = mean(Result, na.rm = T),
      dev = Result - avg
    ) %>%
    nest %>%
    mutate(
      res = purrr::map(data, function(x){

        nval <- nrow(na.omit(x))
        knout <- try(kendallTrendTest(dev ~ Year, x), silent = T)
        outpval <- try(p_ast(knout$p.value), silent = T)
        
        if(!inherits(outpval, 'try-error')){
          
          outest <- round(knout$estimate, 2)
          out <- c(n = nval, outest, pval = outpval) %>%
            data.frame %>%
            t %>%
            data.frame %>%
            select(-intercept) %>% 
            mutate_if(is.factor, function(x) as.character(x))
          
        }
        
        if(inherits(outpval, 'try-error')){
          
          out <- data.frame(n = nval, tau = NA, slope = NA, pval = NA)
          
        }
 
        return(out)
        
      })
    ) %>%
    select(-data) 
  
  res <- do.call('rbind', out$res)
  out <- out %>% 
    select(-res) %>% 
    bind_cols(res) %>% 
    mutate(
      trend = ifelse(tau < 0, 'dec', 'inc'),
      trend = factor(trend, levels = c('inc', 'dec')), 
      slope = as.numeric(slope),
      n = as.numeric(n),
      tau = as.numeric(tau)
    ) %>% 
    ungroup()

  return(out)
  
})

# table for all trends
alltrndtab <- reactive({
  
  # input
  alltrnd <- alltrnd()
  
  totab <- alltrnd
  
  HTML(knitr::kable(totab, format = 'html') %>%
    kable_styling(full_width = T, font_size = 14))

})

# map for all trends
alltrndmap <- reactive({

  # input
  alltrnd <- alltrnd()

  locs <- lddat %>%
    select(StationCode, Longitude, Latitude) %>%
    unique

  tomap <- alltrnd %>%
    left_join(locs, by = 'StationCode') %>%
    na.omit %>% 
    st_as_sf(coords = c("Longitude", "Latitude"), crs = prj)

  # point colors
  cols <- c('tomato1', 'lightgreen')
  names(cols) <- c('dec', 'inc')

  # size values
  cexv <- tomap %>%
    pull(tau) %>%
    abs %>%
    scales::rescale(to = c(2, 15))

   # hover pt labels
  labs <- paste(tomap$StationCode, ': ', tomap$trend, ', tau = ', tomap$tau, ', p = ', tomap$pval)
  mapviewOptions(leafletHeight = 500)

  out <- mapview(tomap, zcol = 'trend', cex = cexv, label = labs, col.regions = cols, legend = F, map.types = mptyps)@map

  return(out)

})

```

This website presents an evaluation of the Orange County mass emissions loading monitoring dataset. The content is separated into four main tabs.  

* __Inventory__: Map-based and tabular summaries of monitoring effort and basic characteristics of the data
* __Trends and power analyses__: Changes over time by select constituents and locations, including power analyses to help identify optimal sampling effort
* __Station differences__: A comparison of time series between stations to identify similarities among trends
* __Overall trends__: Map-based and tabular summary of trend tests for all stations shown together

Each main tab includes sub-tabs or drop-down menus for selecting and viewing different content. Most analyses are also grouped by constituent type as nutrients or metals.  The following shows which constituents are included in each type. 

* __Nutrients__: `r paste(nutrs, collapse = ', ')`
* __Metals__: `r paste(metals, collapse = ', ')`

## Inventory {.tabset .tabset-pills}

### Maps

These maps show the monitoring stations where mass emissions loading data are collected.  The top map shows the watershed for each station and the bottom map shows the number of years for available data.  The maps are linked so that the mouse pointer is synced between the two.

```{r}
tomap <- lddat %>% 
  select(StationCode, Watershed, Longitude, Latitude) %>% 
  unique %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = prj)

m1a <- mapview(tomap, zcol = 'Watershed', layer.name = 'Watershed location', homebutton = F, map.types = mptyps)

tomap <- lddat %>% 
  select(StationCode, yr = Date, Longitude, Latitude) %>% 
  mutate(yr = year(yr)) %>% 
  unique %>% 
  group_by(StationCode, Longitude, Latitude) %>% 
  summarise(n = n()) %>% 
  ungroup %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = prj)

m2a <- mapview(tomap, zcol = 'n', layer.name = 'Number of years', homebutton = F, col.regions = magma, map.types = mptyps)

leafsync::sync(m1a, m2a, ncol = 1) 
```

### Tabular

These tables provide a more detailed summary of the data available at each station.  The drop-down menus can be used to select a constituent type (nutrients or organics) and the type of data shown in each table.  The data type can be selected as the total number of observations, the variance across observations, and the median absolute deviance (MAD) of observations for each station/constituent combination.  The MAD estimates provide a more robust estimate of variance when outliers are present or the data are highly skewed.  

```{r}
column(12, 
       column(4,
         selectInput('consel1', 'Select constituent type:', choices = list('Metals' =  'metals', 'Nutrients' = 'nutrs'), selected = 'nutrs') 
       ),
       column(4, 
         selectInput('tabtyp', 'Table type:', choices = c('Number of observations', 'Variance', 'Median absolute deviation'))
       )
)
```

```{r}
renderUI(invtab())
```

## Trends {.tabset .tabset-pills}

This tab shows trend results for selected constituents. The drop-down menus are filters to select the data to observe. The constituent type and station menus filter which constituents can be selected from the constituent menu. Each sub-tab contains additional information about the results.

```{r}
column(12, 
  column(6, 
       selectInput('consel2', 'Select constituent type:', choices = list('Metals' =  'metals', 'Nutrients' = 'nutrs'), selected = 'nutrs')
  ),
  column(6,
       renderUI({
       
         req(input$varsel)
         
         # input
         varsel <- input$varsel

         sels <- lddat %>% 
           filter(Parameter %in% varsel) %>% 
           pull(StationCode) %>% 
           unique
         
         selectInput('stasel', 'Select station:', choices = sels)  
         
       })
  )
)

column(12,
  column(6, 
       renderUI({
       
         # input
         consel2 <- input$consel2
         
         sels <- eval(parse(text = consel2))
         selectInput('varsel', 'Select constituent:', choices = sels)
         
       })
  )
)

```

### Observed

These plots show the observed time series for `r renderText(input$varsel)` at `r renderText(input$stasel)`. The top plot is the raw data with load estimates in the first row and flow volume in the second row. This plot indicates trends in load over time and if load was correlated with flow volume. The middle plot shows boxplots of the raw data grouped by year, and the bottom plot shows boxplots of the raw data grouped by month.

```{r}
renderPlot({obsp1()}, height = 500, width = 700)
renderPlot({obsp2()}, height = 300, width = 700)
renderPlot({obsp3()}, height = 300, width = 700)
```

### Trends

This sub-tab shows the annual trends for `r renderText(input$varsel)` at `r renderText(input$stasel)`.  The top plot shows trends for the time series after removing the overall mean. This is an effective approach for trend analysis by evaluating deviation of the annual averages relative to the grand mean. Bars in green show annual averages greater than the grand mean and bars in red show annual averages less than the grand mean.  The bars are centered on the zero line, which is a reference for the grand mean. The y-axis reports the grand mean of the constituent in parentheses. The trend is shown as the dotted line for a regression of the annual averages over time.  

```{r}
renderPlot({trndp1()}, height = 350, width = 900)
```

This table shows the regression results from the above plot.  Parameter estimates (Year and Intercept) are shown with their standard errors in parentheses.  

```{r, results='asis'}
renderUI({trndtab1()})
```

<br>

Finally, Kendall tests provide an alternative indication of trend by evaluating magnitude, direction, and significance of a change over time. The test is a non-parametric equivalent to the regression analysis above. The value for tau ranges from -1 to 1 and provides a measure of trend direction. The slope is the estimated change per year in density and the p-value shows the significance of the test.

```{r, results = 'asis'}
renderUI({trndtab2()})
```

## Station differences {.tabset .tabset-pills}

This tab shows similarities among stations for selected constituents.  This information can be useful to identify which stations have similar loading characteristics over time and which stations do not.  Sampling less frequently at stations with similar characteristics may be useful for optimizing efficiency of the monitoring program.  

```{r, fig.height = 5, fig.width = 6, fig.align = 'center'}
# rename some nutrient parameters
ldprp <- lddat %>%
  mutate(
    Date = floor_date(Date, unit = 'month')
  ) %>%
  filter(Parameter %in% c(metals, nutrs)) %>%
  select(Watershed, StationCode, Date, Parameter, Result) %>%
  group_by(Watershed, StationCode, Date, Parameter) %>%
  summarise(Result = mean(Result, na.rm = T)) %>%
  group_by(StationCode, Parameter) %>%
  mutate(Result = case_when(
    is.na(Result) ~ mean(Result, na.rm = T),
    T ~ Result
    )
  ) %>%
  ungroup()

siteshd <- ldprp %>%
  select(Watershed, StationCode) %>%
  unique
```

```{r}
column(12,
  column(6,
        selectInput('consel3', 'Select constituent type:', choices = list('Metals' =  'metals', 'Nutrients' = 'nutrs'), selected = 'nutrs')
         ),
  column(6,
         renderUI({

           # input
           consel3 <- input$consel3

           sels <- eval(parse(text = consel3))

           selectInput('varsel2', 'Select constituent:', choices = sels)

         })
  )
)
column(12,
  column(6,
       selectInput('stasel2', 'Select first station:', choices = sort(unique(lddat$StationCode)), selected = 'SDMF05')
  ),
  column(6,
       selectInput('stasel3', 'Select second station:', choices = sort(unique(lddat$StationCode)), selected = 'SADF01')
  )
)
```

This plot shows a Principal Components Analysis (PCA) comparison of all sites, dates, and constituents and a dissimilarity matrix for `r renderText({input$varsel2})` between the time series at each station.  Blue tiles are more similar and red are more dissimilar.  The larger points (black and transparent black) in the PCA plot correspond to the observations at the two selected stations from the drop-down menus above.  The dissimilarity estimate for the two selected sites is also emphasized by black outline in the right plot. Sites with points that closely overlap on the PCA plot have lower dissimilarity measures (blue tiles), whereas sites with less overlap show higher dissimilarity measures (red tiles).  

```{r}
renderPlot({mulplo()}, height = 550, width = 900)
```

Information for the two selected stations above can be further evaluated by viewing the observed time series.  Sites that are similar will have time series with similar characteristics.  This will likely apply to the observed data, the boxplots by year, and the boxplots by month. 

### Time series

```{r}
renderPlot({obsp4()}, height = 450, width = 700)
```

### By year

```{r}
renderPlot({obsp5()}, height = 450, width = 700)
```

#### By month

```{r}
renderPlot({obsp6()}, height = 450, width = 700)
```

## Overall trends

```{r}
column(12, 
  column(6, 
       selectInput('consel4', 'Select constituent type:', choices = list('Metals' =  'metals', 'Nutrients' = 'nutrs'),  selected = 'nutrs')
  ),
  column(6, 
       renderUI({
       
         # input
         consel4 <- input$consel4

         sels <- eval(parse(text = consel4))
         selectInput('varsel3', 'Select constituent:', choices = sels)
         
       })
  )
)

```

This final tab shows results of trends at all stations using Kendall tests. These tests provide a non-parametric indication of a trend by evaluating magnitude, direction, and significance of a change over time. The value for tau ranges from -1 to 1 and provides a measure of trend direction. The slope is the estimated change per year in density and the p-value shows the significance of the test.

The maps show the value for tau (direction of trend) for a Kendall test for changes across stations, green for decreasing and red for increasing. Size of the point is the magnitude of the estimated change over time.  The table shows the detailed results from the map.

```{r}
renderLeaflet(alltrndmap())
```

```{r}
renderUI(alltrndtab())
```

